Scaling rates of true polar wander in convecting planets and moons

### 2017-09-11 tpw_rate.tex
Wordcrafting.
Stronger statement.

\begin{equation}
\frac{d \mathbf{H}}{dt} + \mitbf{\Omega} \times \mathbf{H} = 0
\label{eq:euler}
\end{equation}
where $\mathbf{H} = \mathbf{I} \cdot \mitbf{\Omega}$ is the angular momentum vector, $\mathbf{I}$ is the moment of inertia tensor, and $\mitbf{\Omega}$ is the angular velocity vector.
On a dynamic planet $\mathbf{I}$ may be a function of time, so $\mitbf{\Omega}$ must be also vary with time to conserve angular momentum.
In this case, Equation~\eqref{eq:euler} is often called the Liouville equation \citep[e.g.][]{munk1960rotation}.
For a slowly convecting fluid, such as a planetary mantle, the inertial term $\partial \mathbf{H} / \partial t$ is negligible, so we may solve the simplified quasistatic equations
\begin{equation}
\mitbf{\Omega } (t)\times ( \mathbf{I}(t) \cdot \mitbf{\Omega }(t) ) = 0.
\label{eq:quasistatic_liouville}
\end{equation}
Equation~\eqref{eq:quasistatic_liouville} indicates that $\mitbf{\Omega}$ and $\mathbf{H}$ are parallel, so a solution for $\mitbf{\Omega}(t)$ is equivalent to solving an eigenvalue problem for the principal values of $\mathbf{I}(t)$;  the eigenvectors define the principal axes and the eigenvalues correspond to the principal moments of inertia. The rotation axis is expected to align with the axis for the largest moment.  
In practice, this eigenvalue approach has been often used in previous studies for computing the spin axis of a planet \citep[e.g.][]{steinberger1997changes, roberts2007cause}.
The moment of inertia tensor in Equation~\eqref{eq:quasistatic_liouville} includes all contributions to the mass structure
of the planet, including the spherically symmetric mass distribution, rotational deformation, deformation due to self gravity, internal and surface density anomalies, and the associated surface deflections.
Here we are interested in those contributions that arise from mantle convection.

@@

The infinite-time limit of Equation~\eqref{eq:viscoelastic_deformation} for a 
constant rotation vector around the $z$-axis implies
\begin{equation}
\begin{aligned}
J_{zz} = C = &\frac{2}{3} \frac{k_f a^5 \Omega^2}{3 G} \\
J_{xx} = J_{yy} = A = -&\frac{1}{3} \frac{k_f a^5 \Omega^2}{3 G} \\
\end{aligned}
\end{equation}
where $C$ and $A$ are the polar and equatorial moments of inertia, respectively, 
and $k_f$ is the fluid limit of $k$. 
We can solve for $k_f$ in terms of $C-A$:
\begin{equation}
k_f = \frac{3 G (C-A)}{\Omega^2 a^5}.
\label{eq:fluid_love}
\end{equation}
This infinite-time limit  ensures that the left-hand side of Equation (4) vanishes. It follows that the contribution from the convective perturbation $E_{ij}$ also vanishes, so TPW is not permitted.

>>>>

\begin{equation}
\frac{d \mathbf{H}}{dt} + \mitbf{\Omega} \times \mathbf{H} = 0
\label{eq:euler}
\end{equation}
where $\mathbf{H} = \mathbf{I} \cdot \mitbf{\Omega}$ is the angular momentum vector, $\mathbf{I}$ is the moment of inertia tensor, and $\mitbf{\Omega}$ is the angular velocity vector.
On a dynamic planet $\mathbf{I}$ is a function of time, so $\mitbf{\Omega}$ must also vary with time to conserve angular momentum.
In this case, Equation~\eqref{eq:euler} is often called the Liouville equation \citep[e.g.][]{munk1960rotation}.
For a slowly convecting fluid, such as a planetary mantle, the inertial term $\partial \mathbf{H} / \partial t$ is negligible, so we may solve the simplified quasistatic equations
\begin{equation}
\mitbf{\Omega } (t)\times ( \mathbf{I}(t) \cdot \mitbf{\Omega }(t) ) = 0.
\label{eq:quasistatic_liouville}
\end{equation}
Equation~\eqref{eq:quasistatic_liouville} indicates that $\mitbf{\Omega}$ and $\mathbf{H}$ are parallel, so a solution for $\mitbf{\Omega}(t)$ is equivalent to solving an eigenvalue problem for the principal values of $\mathbf{I}(t)$;  the eigenvectors define the principal axes and the eigenvalues correspond to the principal moments of inertia. The rotation axis is expected to align with the axis for the largest moment.  
In practice, this eigenvalue approach has been often used in previous studies for computing the spin axis of a planet \citep[e.g.][]{steinberger1997changes, roberts2007cause}.
The moment of inertia tensor in Equation~\eqref{eq:quasistatic_liouville} includes all contributions to the mass structure
of the planet, including the spherically symmetric mass distribution, rotational deformation, deformation due to self gravity, internal and surface density anomalies, and the associated surface deflections.
Here we are interested in those contributions that arise from mantle convection.

@@

The infinite-time limit of Equation~\eqref{eq:viscoelastic_deformation} for a 
constant rotation vector around the $z$-axis implies
\begin{equation}
\begin{aligned}
J_{zz} = C = &\frac{2}{3} \frac{k_f a^5 \Omega^2}{3 G} \\
J_{xx} = J_{yy} = A = -&\frac{1}{3} \frac{k_f a^5 \Omega^2}{3 G} \\
\end{aligned}
\end{equation}
where $C$ and $A$ are the polar and equatorial moments of inertia, respectively, 
and $k_f$ is the fluid limit of $k$. 
We can solve for $k_f$ in terms of $C-A$:
\begin{equation}
k_f = \frac{3 G (C-A)}{\Omega^2 a^5}.
\label{eq:fluid_love}
\end{equation}
This infinite-time limit  ensures that the left-hand side of Equation (4) vanishes. It follows that the contribution from the convective perturbation $E_{ij}$ also vanishes, so no further TPW is permitted.
